Phonetic Tansfer analysis code

### 2022-05-19 RCode/transfer_analysis.Rmd
manuscript revision: major edits, figures

library(tidyverse)
library(phonR)
library(lme4)
library(stargazer)
library(modelsummary)
library(lmerTest)
library(pwr)
library(simr)
library(EMAtools) # to calculate Cohen's d values from lmer objects

@@

# replacing vowel symbols with unicode:
remapping_english <- c(`\\ae` = "æ", `\\vt` = "ʌ")

@@

# data frame with time as variable

@@

# raw FFs
df_time <- df %>%
  select(!c("filename", "f1_mid", "vowel_duration")) %>%
  rename(f1_1=f1_5, f1_2=f1_15, f1_3=f1_25, f1_4=f1_35, f1_5=f1_45, f2_1=f2_5, f2_2=f2_15, f2_3=f2_25, f2_4=f2_35, f2_5=f2_45) %>%
  # dplyr pivot() functions: convert data between long(indexed) and wide(cartesian) formats
  pivot_longer(starts_with("f"),
               names_to = c(".value", "time"),
               names_sep = "_",
               names_transform = list(time = as.factor)
               )

@@

@@

# IPA vowel symbols
remapping_english <- c(`\\ae` = "æ", `\\vt` = "ʌ")
remapping_bengali <- c(A = "a:", E = "æ")

# add column with unicode vowels
df_ben_normalized$unicodevowel <- remapping_bengali[as.character(df_ben_normalized$vowel)]

df_normalized$unicodevowel <- remapping_english[as.character(df_normalized$vowel)]

# plot: unilingual English, mixed, and baseline Bengali
xlim <- c(2500,600)
ylim = c(1100,400)

temp <- df_ben %>%
  filter(time %in% (8:12), f1>300)

with(temp,plotVowels(f1, f2, vowel,
                        plot.tokens = F, plot.means = TRUE, pch.tokens = vowel, pch.means = vowel, 
                        var.col.by = vowel, var.sty.by = vowel, 
                        cex.tokens = 0.8, cex.means = 1, alpha.means = 0.8,alpha.tokens = 0.3,
                        ellipse.line = F, ellipse.fill = T, fill.opacity = 0.3,
                        poly.line = F, poly.order = c("A","E"),
                        xlim = xlim, ylim = ylim, 
                        legend.kwd="bottomleft", legend.args = list(legend=c("baseline æ","baseline a:")), 
                        axes=T, pretty = TRUE))

# Bengali vowels
with(df_ben_normalized,plotVowels(f1_raw, f2_raw, vowel,
                        plot.tokens = F, plot.means = TRUE, pch.tokens = unicodevowel, pch.means = unicodevowel, 
                        var.col.by = vowel, var.sty.by = vowel, 
                        cex.tokens = 0.8, cex.means = 1, alpha.means = 0.8,alpha.tokens = 0.3,
                        ellipse.line = F, ellipse.fill = T, fill.opacity = 0.2,
                        poly.line = F, poly.order = c("A","E"),
                        xlim = xlim, ylim = ylim, 
                        legend.kwd="bottomleft", legend.args = list(legend=c("baseline æ","baseline a:")), 
                        axes=FALSE, pretty = TRUE))
par(new=TRUE)

# English vowels
with(df_normalized, plotVowels(f1_raw, f2_raw, vowel, group = context, 
                               plot.tokens = F, plot.means = TRUE , pch.tokens = unicodevowel, pch.means = unicodevowel,
                               var.col.by = NA, var.sty.by = context, 
                               cex.tokens = 0.8, cex.means = 1, alpha.means = 0.5,alpha.tokens = 0.3,
                               ellipse.line = T,  ellipse.fill = T,
                               poly.line = F, poly.order = c("\\ae","\\vt"),
                               xlim = xlim, ylim = ylim, 
                               legend.kwd = "bottomright", legend.args = list(legend=c("unilingual","mixed")),
                               fill.opacity = 0.1, pretty = TRUE))

@@

```{r}
# plot for English vowels:

@@

```{r, results='hide'}

@@

summary(phon_acco.model_f1)
summary(rePCA(phon_acco.model))

@@

stargazer(phon_acco.model_f1)

@@

anova(phon_acco.model, null.model)

# some weird interaction with lmerTest package. Detach that first
stargazer(phon_acco.model)
modelsummary(phon_acco.model)

@@

```{r}
# F2

# full model
phon_acco.model_f2 = lmer(f2 ~ 
                         vowel*time + vowel*context*task + task*time +
                         (1+context|word) + (1+context+vowel|subject), 
                         data= temp, REML=FALSE, control = lmerControl(optimizer = "bobyqa"))
summary(phon_acco.model_f2)
summary(rePCA(phon_acco.model))

@@

anova(phon_acco.model, null.model)

stargazer(phon_acco.model)
```

@@

@@

# setting beta-values to 85% of original
beta.f1<- beta.f1 %>%
  map(~ . * 0.95) %>% # map each element of the vector to the result of applying the function f to it (creates a list)
  unlist()  # convert back to a named vector

@@

power_f1_minus5 <- mixedpower(model = phon_acco.model_f1, data = temp,
                          fixed_effects = c("vowel","context","time","task"),
                          simvar = "subject_num", steps = c(5,8,10,12,15), n_sim = 500,
                          SESOI = beta.f1, databased = F,
                          critical_value = 2)

powerplot_f1_minus5 <- multiplotPower(power_f1_minus5, filename = "powerplot_f1_minus5.png")

@@

# setting beta-values to 85% of original
beta.f2 <- beta.f2 %>%
  map(~ . * 1.15) %>% # map each element of the vector to the result of applying the function f to it (creates a list)
  unlist()  # convert back to a named vector

@@

power_f2_plus15 <- mixedpower(model = phon_acco.model_f2, data = temp,
                          fixed_effects = c("vowel","context","time","task"),
                          simvar = "subject_num", steps = c(5,7,10,12,15), n_sim = 500,
                          SESOI = beta.f2, databased = F,
                          critical_value = 2)

powerplot_f2_plus15 <- multiplotPower(power_f2_plus15, filename = "powerplot_f2_plus15.png")

@@

```{r}
# scratch pad: additional code for manuscript revision

>>>>

library(tidyverse)
library(phonR)
library(lme4)
#library(stargazer)
#library(modelsummary)
library(lmerTest)
#library(pwr)
#library(simr)
#library(EMAtools) # to calculate Cohen's d values from lmer objects

@@

# for replacing vowel symbols with unicode:
remapping_english <- c(`\\ae` = "æ", `\\vt` = "ʌ")
remapping_bengali <- c(A = "a:", E = "æ")

@@

# data frame with time as variable (long format)

@@

# raw FFs with time as variable
df_time <- df %>%
  select(!c("filename", "f1_mid", "vowel_duration")) %>%
  rename(f1_1=f1_5, f1_2=f1_15, f1_3=f1_25, f1_4=f1_35, f1_5=f1_45, f2_1=f2_5, f2_2=f2_15, f2_3=f2_25, f2_4=f2_35, f2_5=f2_45) %>%
  # dplyr pivot() functions: convert data between long(indexed) and wide(cartesian) formats
  pivot_longer(starts_with("f"),
               names_to = c(".value", "time"),
               names_sep = "_",
               names_transform = list(time = as.factor)
               )

@@

df_ben_AE <- df_ben %>%
  filter(grepl("A|E", vowel), time %in% (8:12), f1>600) %>%
  droplevels() %>%
  mutate(time_absolute = NULL) %>%
  relocate(gender, .after = speaker) %>%
  mutate (vowel = remapping_bengali[as.character(vowel)])

@@

# add column with unicode vowels (if needed)
#df_normalized$unicodevowel <- remapping_english[as.character(df_normalized$vowel)]

# plot: unilingual English, mixed, and baseline Bengali
xlim <- c(2500,600)
ylim <- c(1100,500)

## plot for baseline Bengali vs unilingual English
df_eng_unilingual <- df_time %>%
  filter(time %in% (3:5), context == "e")

with(df_ben_AE,plotVowels(f1, f2, vowel,
                        plot.tokens = F, plot.means = TRUE, pch.tokens = vowel, pch.means = vowel, 
                        var.col.by = NULL, var.sty.by = NULL, 
                        cex.tokens = 0.8, cex.means = 1, alpha.means = 0.8,alpha.tokens = 0.3,
                        ellipse.line = F, ellipse.fill = T, fill.opacity = 0.2, 
                        poly.line = F, poly.order = c("a:","æ"),
                        xlim = xlim, ylim = ylim, 
                        #legend.kwd="bottomleft", legend.args = list(legend=c("baseline æ","baseline a:")), 
                        pretty = TRUE))
par(new=TRUE)

# English vowels
with(df_eng_unilingual, plotVowels(f1, f2, vowel, 
                               plot.tokens = F, plot.means = TRUE , pch.tokens = vowel, pch.means = vowel,
                               var.col.by = NULL, var.sty.by = NULL, 
                               cex.tokens = 0.8, cex.means = 1, alpha.means = 0.5,alpha.tokens = 0.3,
                               ellipse.line = T, ellipse.fill = F, fill.opacity = 0.1,
                               poly.line = F, poly.order = c("æ","ʌ"),
                               xlim = xlim, ylim = ylim, 
                               #legend.kwd = "bottomright", 
                               axes=FALSE, pretty = TRUE))

rm(df_eng_unilingual)

```{r}
## final result plot:

df_eng_plots <- df_time %>%
  filter(time %in% (3:5))

# Bengali vowels
with(df_ben_AE,plotVowels(f1, f2, vowel,
                        plot.tokens = F, plot.means = TRUE, pch.tokens = vowel, pch.means = vowel, 
                        var.col.by = vowel, var.sty.by = vowel, 
                        cex.tokens = 0.8, cex.means = 1, alpha.means = 0.8,alpha.tokens = 0.3,
                        ellipse.line = F, ellipse.fill = T, fill.opacity = 0.2,
                        poly.line = F, poly.order = c("a:","æ"),
                        xlim = xlim, ylim = ylim, 
                        legend.kwd="bottomleft", legend.args = list(legend=c("baseline æ","baseline a:")), 
                        pretty = TRUE))
par(new=TRUE)

# English vowels
with(df_eng_plots, plotVowels(f1, f2, vowel, group = context,
                               plot.tokens = F, plot.means = TRUE , pch.tokens = vowel, pch.means = vowel,
                               var.col.by = NULL, var.sty.by = context, 
                               cex.tokens = 0.8, cex.means = 1, alpha.means = 0.5,alpha.tokens = 0.3,
                               ellipse.line = T, ellipse.fill = T, fill.opacity = 0.1,
                               poly.line = F, poly.order = c("æ","ʌ"),
                               xlim = xlim, ylim = ylim, 
                               legend.kwd = "bottomright", legend.args = list(legend=c("unilingual","mixed")),
                               axes=FALSE, pretty = TRUE))

rm(df_eng_plots)

@@

```{r, eval=FALSE, include=FALSE}
# plot for English vowels:

@@

@@

summary(phon_acco.model_f1)
summary(rePCA(phon_acco.model_f1))

@@

#stargazer(phon_acco.model_f1)

@@

anova(phon_acco.model_f1, null.model)

# some weird interaction with lmerTest package. Detach that first
#stargazer(phon_acco.model)
#modelsummary(phon_acco.model)

@@

```{r,eval=FALSE}
# F2

# full model
phon_acco.model_f2 = lmer(f2 ~ 
                         vowel*time + vowel*context*task + task*time +
                         (1+context|word) + (1+context+vowel|subject), 
                         data= temp, REML=FALSE, control = lmerControl(optimizer = "bobyqa"))
summary(phon_acco.model_f2)
summary(rePCA(phon_acco.model_f2))

@@

anova(phon_acco.model_f2, null.model)

@@

```{r, include=FALSE, eval=FALSE}

@@

# setting beta-values to 85% of original
beta.f1<- beta.f1 %>%
  map(~ . * 1.05) %>% # map each element of the vector to the result of applying the function f to it (creates a list)
  unlist()  # convert back to a named vector

@@

power_f1_plus5 <- mixedpower(model = phon_acco.model_f1, data = temp,
                          fixed_effects = c("vowel","context","time","task"),
                          simvar = "subject_num", steps = c(5,8,10,12,15), n_sim = 500,
                          SESOI = beta.f1, databased = F,
                          critical_value = 2)

powerplot_f1_plus5 <- multiplotPower(power_f1_plus5, filename = "powerplot_f1_plus5.png")

@@

# setting beta-values to 85% of original
beta.f2 <- beta.f2 %>%
  map(~ . * 1.05) %>% # map each element of the vector to the result of applying the function f to it (creates a list)
  unlist()  # convert back to a named vector

@@

power_f2_plus5 <- mixedpower(model = phon_acco.model_f2, data = temp,
                          fixed_effects = c("vowel","context","time","task"),
                          simvar = "subject_num", steps = c(5,8,10,12,15), n_sim = 500,
                          SESOI = beta.f2, databased = F,
                          critical_value = 2)

 powerplot_f2_plus5 <- multiplotPower(power_f2_plus5, filename = "powerplot_f2_plus5.png")

@@

```{r, eval=FALSE, include=FALSE}
# scratch pad: additional code for manuscript revision
